---
title: "Lecture 6"
author: "Michal Kubišta"
date: "2018/10/31"
output:
  xaringan::moon_reader:
    lib_dir: libs
    chakra: ../remark.js
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

## Structure
- Conditionals
- Functions
- Loops & vectorisation
```{r setup, include = FALSE}
library(data.table)
library(tidyverse)
```

---
class: middle, inverse

# Conditionals

---

class: middle, inverse

# Functions

---

class: middle, inverse

# Loops & vectorisation

---

## Doing operations again & again

```{r loop_example}
mtcars$mpg = as.numeric(mtcars$mpg)
mtcars$cyl = as.numeric(mtcars$cyl)
mtcars$disp = as.numeric(mtcars$disp)
# dont
```

---

## Loops
- do the same thing with different input

```{r loop_example2}
cols = c("mpg", "cyl", "disp")

for (i in cols) {
    mtcars[,i] = as.numeric(mtcars[,i])
}
```

---

## Type of loops

.pull-left[
- **for loop**
    - for every element, do
```{r for_loop}
for (i in 1:4) {
    print(i)
}
```
]
.pull-right[
- **while**
    - while condition true, do something
```{r while_loop}
test = 8
while (test < 10) {
    print(test)
    test = test + 1
}
```
]

---

## Pros & Cons
.pull-left[
PRO
- still better than code repetition
- easy to comprehend
]

.pull-right[
CON
- TERRIBLE performance
]

--------------------------------------------   

```{r loop_vec comparison}
a = seq_len(100000000)
system.time({
    for (i in a) {
        i^2
    }
})
system.time({b = data.table::data.table(a)[, a^2]})
```

```{r cleanup, include = FALSE}
rm(a, b)
```

---

## So.....?
- loops are useful for
    - small number of repetitions
    - when the number of repetitions is unknown (simulations)

## For the rest there's vectorisation!
- apply the function to each element
- heavily optimised

```{r vectorise}
map_dbl(1:20, ~.x^2)
map_dbl(1:20, sqrt)
```

