---
title: "Lecture 5"
author: "Michal Kubišta"
date: "2018/10/24"
output:
  xaringan::moon_reader:
    lib_dir: libs
    chakra: ../remark.js
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

## Structure
- Casting
- Binding
- Joins
```{r setup, include = FALSE}
library(tidyverse)
```

---
class: middle, inverse

# Casting

---

## Gather
- collect several columns into key-value based (two) columns

```{r gather_setup}
cars = 
    mtcars[1:5,] %>% 
    mutate(name = rownames(.)) %>% 
    select(name, mpg, cyl, disp)
cars

```

---

## Gather
- collect several columns into key-value based (two) columns

```{r gather}
cars_g = 
    cars %>% 
    gather(fact, value, mpg, cyl, disp)
cars_g
```

---

## Spread
- inverse of gather
- spread key-value columns into several columns

```{r spread}
cars_g %>% 
    spread(fact, value)
```

---
class: middle, inverse

# Binding

---

## Rbind
- bind rows
- two tables of the same variables

```{r rbind_setup}
(cars_a = cars[1:3,])
(cars_b = cars[4:5,])
```

---

## Rbind
- glue rows of tables
- two tables of the same variables (same columns)

```{r rbind}
rbind(cars_a, cars_b)
```

---

## Cbind
- glue columns of tables
- two tables of the same observation (same row)

```{r cbind_setup}
(cars_a = cars %>% select(name, mpg))
(cars_b = cars %>% select(cyl, disp))
```

---

## Cbind
- glue columns of tables
- two tables of the same observation (same row)

```{r cbind}
cbind(cars_a, cars_b)
```

---

class: middle, inverse

# Joins

---
